---
title: "CWNS Wastewater Treatment Plant Location Correction"
subtitle: "Two-Stage ML Pipeline — Results Review"
author: "Andrew Murray"
date: today

# ── Version parameter ──────────────────────────────────────────────────────────
# By default (version: null) the report auto-detects the highest-numbered
# Version_X folder under Location_Correction/results/.
# Override with an explicit string to re-run an older version, e.g.:
#   quarto render cwns_location_correction_report.qmd -P version:Version_1
params:
  version: null   # null = auto-detect highest Version_X; "Version_N" to override

format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-tools: true
    fig-width: 9
    fig-height: 5.5
    fig-dpi: 150
    embed-resources: true
    df-print: paged
    css: styles.css
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

```{r setup}
#| cache: false
library(tidyverse)
library(arrow)
library(here)
library(sf)
library(gt)
library(gtExtras)
library(scales)
library(patchwork)
library(ggridges)
library(glue)

# ── Version resolution ─────────────────────────────────────────────────────────
# If params$version is NULL, auto-detect the highest-numbered Version_X folder.
# Override by rendering with: quarto render ... -P version:Version_1
.versions_root <- here("Location_Correction", "results")

if (is.null(params$version)) {
  .all_dirs  <- list.dirs(.versions_root, recursive = FALSE, full.names = FALSE)
  .ver_dirs  <- .all_dirs[grepl("^Version_[0-9]+$", .all_dirs)]
  if (length(.ver_dirs) == 0) {
    stop("No Version_N folders found under: ", .versions_root)
  }
  .ver_nums     <- as.integer(sub("Version_", "", .ver_dirs))
  version_label <- .ver_dirs[which.max(.ver_nums)]
} else {
  version_label <- params$version
}

message("Using model version: ", version_label)


# ── Paths ──────────────────────────────────────────────────────────────────────
results_dir <- file.path(.versions_root, version_label)

path_stage1  <- file.path(results_dir, "stage1_results.parquet")
path_stage2  <- file.path(results_dir, "stage2_results.parquet")
path_summary <- file.path(results_dir, "plant_summary.csv")
path_gpkg    <- file.path(results_dir, "inference_results.gpkg")

# ── Thresholds (from training) ─────────────────────────────────────────────────
STAGE1_THRESHOLD <- 0.677
STAGE2_THRESHOLD <- 0.057

# ── Theme ──────────────────────────────────────────────────────────────────────
epa_blue   <- "#005ea2"
epa_green  <- "#2e8540"
epa_red    <- "#d83933"
epa_gold   <- "#f5a623"
epa_grey   <- "#71767a"
epa_light  <- "#f0f4f9"

theme_cwns <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title        = element_text(face = "bold", colour = epa_blue, size = base_size + 2),
      plot.subtitle     = element_text(colour = epa_grey, size = base_size - 1),
      plot.caption      = element_text(colour = epa_grey, size = base_size - 3, hjust = 0),
      axis.title        = element_text(colour = epa_grey, size = base_size - 1),
      strip.text        = element_text(face = "bold", colour = epa_blue),
      panel.grid.minor  = element_blank(),
      panel.grid.major  = element_line(colour = "#e5e8ec"),
      plot.background   = element_rect(fill = "white", colour = NA),
      panel.background  = element_rect(fill = "white", colour = NA),
      legend.position   = "bottom"
    )
}

theme_set(theme_cwns())

# ── Tier colour/label lookup ───────────────────────────────────────────────────
# tier_label is already present in the summary CSV.
# Define ordered factor levels and a matching colour palette for consistency.
tier_levels <- c(
  "T1 — Nearby correction (<5 km)",
  "T1 — Distant correction (>=5 km)",
  "T1 — No candidate found",
  "T2 — No parcel, corrected",
  "T2 — No parcel, uncorrected",
  "T3 — Likely correct"
)

tier_colors <- c(
  "T1 — Nearby correction (<5 km)"   = epa_red,
  "T1 — Distant correction (>=5 km)" = "#a83228",
  "T1 — No candidate found"          = "#e07b7b",
  "T2 — No parcel, corrected"        = epa_gold,
  "T2 — No parcel, uncorrected"      = "#c8860a",
  "T3 — Likely correct"              = epa_green
)

# Raw tier values as they appear in the 'tier' column
tier_raw_corrected <- c(
  "Tier1_nearby_correction",
  "Tier1_distant_correction",
  "Tier2_no_parcel_corrected"
)
```

```{r load-data}
#| cache: true

# ── Load datasets ──────────────────────────────────────────────────────────────
# Stage 1 columns:
#   stage1_prob_correct, stage1_prob_incorrect, CWNS_ID, no_parcel_found,
#   trigger_reason
#
# Stage 2 columns:
#   stage2_prob_correct, stage2_prob_incorrect, CWNS_ID, ll_uuid, h3_index_9,
#   state, distance_m, within_1km, within_5km
#
# Summary columns (all of the above, plus):
#   top_stage2_prob, top_ll_uuid, top_distance_m, within_1km, within_5km,
#   FACILITY_NAME, LATITUDE, LONGITUDE, STATE_CODE, pop_served, requires_npdes,
#   surface_water_discharge, subdivision, place, county, is_rural, tier

stage1     <- read_parquet(path_stage1)
stage2     <- read_parquet(path_stage2)
summary_df <- read_csv(path_summary, show_col_types = FALSE)

# ── Derive tier_label from tier ───────────────────────────────────────────────
tier_map <- c(
  "Tier1_nearby_correction"     = "T1 — Nearby correction (<5 km)",
  "Tier1_distant_correction"    = "T1 — Distant correction (>=5 km)",
  "Tier1_no_candidate"          = "T1 — No candidate found",
  "Tier2_no_parcel_corrected"   = "T2 — No parcel, corrected",
  "Tier2_no_parcel_uncorrected" = "T2 — No parcel, uncorrected",
  "Tier3_likely_correct"        = "T3 — Likely correct"
)

summary_df <- summary_df |>
  mutate(tier_label = factor(recode(tier, !!!tier_map), levels = tier_levels))

# ── Derived counts ─────────────────────────────────────────────────────────────
n_total   <- nrow(stage1)
n_flagged <- sum(stage1$stage1_prob_correct < STAGE1_THRESHOLD | stage1$no_parcel_found)
n_correct <- n_total - n_flagged
n_stage2  <- n_distinct(stage2$CWNS_ID)
n_cands   <- nrow(stage2)

n_corrected_total <- summary_df |>
  filter(!is.na(top_stage2_prob), top_stage2_prob >= STAGE2_THRESHOLD) |>
  nrow()
```

---

# Executive Summary {.unnumbered}

::: {.callout-note appearance="minimal"}
**Model version:** `r version_label` &emsp; | &emsp; **Results directory:** `r results_dir`
:::

This report reviews results from a two-stage machine learning pipeline designed
to detect and correct erroneous reported locations for U.S. wastewater treatment
plants in the **Clean Watersheds Needs Survey (CWNS)** dataset. The pipeline was
trained and run on EPA's HPC cluster (RedHat 8.10, SLURM) using R 4.4.2.

```{r exec-summary-table}
#| echo: false
tibble(
  Metric = c(
    "Plants evaluated (Stage 1)",
    "Plants flagged for Stage 2",
    "Candidate parcels evaluated (Stage 2)",
    "Plants with accepted correction",
    "Stage 1 ROC AUC",
    "Stage 2 ROC AUC",
    "Stage 1 decision threshold (prob_correct)",
    "Stage 2 decision threshold (prob_correct)"
  ),
  Value = c(
    comma(n_total),
    glue("{comma(n_flagged)} ({percent(n_flagged / n_total, 0.1)})"),
    comma(n_cands),
    glue("{comma(n_corrected_total)} ({percent(n_corrected_total / n_total, 0.1)})"),
    "0.986",
    "0.975",
    "0.677",
    "0.057"
  )
) |>
  gt() |>
  tab_header(title = "Pipeline Performance Overview") |>
  cols_label(Metric = "Metric", Value = "Value") |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = epa_light),
    locations = cells_body(rows = seq(1, 8, 2))
  ) |>
  opt_table_font(font = google_font("IBM Plex Sans")) |>
  tab_options(table.width = pct(75))
```

---

# Stage 1 — Location Confidence Scoring

## Score Distribution

Stage 1 assigns each plant a `stage1_prob_correct` — the probability that its
reported location falls on the correct land parcel. The decision threshold of
**0.677** was selected at training to maximise the F-measure on held-out
validation data. Plants are also flagged if `no_parcel_found = TRUE` regardless
of score.

```{r stage1-hist}
#| fig-cap: "Distribution of `stage1_prob_correct` for all CWNS plants. The red
#|   dashed line marks the classification threshold (0.677). Plants left of this
#|   line, or with no parcel found, are sent to Stage 2."

stage1 |>
  mutate(decision = if_else(
    stage1_prob_correct >= STAGE1_THRESHOLD & !no_parcel_found,
    "Likely correct", "Flagged"
  )) |>
  ggplot(aes(x = stage1_prob_correct, fill = decision)) +
  geom_histogram(bins = 80, colour = "white", linewidth = 0.15) +
  geom_vline(xintercept = STAGE1_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.9) +
  annotate("text", x = STAGE1_THRESHOLD + 0.015, y = Inf,
           label = glue("Threshold\n{STAGE1_THRESHOLD}"),
           vjust = 1.3, hjust = 0, size = 3.2, colour = epa_red) +
  scale_fill_manual(values = c("Flagged" = epa_red, "Likely correct" = epa_green)) +
  scale_x_continuous(labels = percent_format(1), breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(labels = comma) +
  labs(
    title   = "Stage 1 Score Distribution",
    subtitle = glue("{comma(n_correct)} plants above threshold · {comma(n_flagged)} flagged"),
    x       = "stage1_prob_correct",
    y       = "Number of plants",
    fill    = NULL,
    caption = "Note: plants with no_parcel_found = TRUE are flagged regardless of score."
  )
```

## Score by Parcel Found Status

```{r stage1-parcel}
#| fig-cap: "Stage 1 score distributions split by `no_parcel_found`. Non-intersecting
#|   plants are sent to Stage 2 regardless of score."

stage1 |>
  mutate(parcel_status = if_else(no_parcel_found, "No parcel found", "Parcel found")) |>
  ggplot(aes(x = stage1_prob_correct, y = parcel_status, fill = parcel_status)) +
  geom_density_ridges(alpha = 0.75, quantile_lines = TRUE,
                      quantiles = 0.5, colour = "white") +
  geom_vline(xintercept = STAGE1_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(values = c("Parcel found" = epa_blue, "No parcel found" = epa_grey),
                    guide = "none") +
  scale_x_continuous(labels = percent_format(1), limits = c(0, 1)) +
  labs(
    title = "Stage 1 Scores by Parcel Found Status",
    x     = "stage1_prob_correct",
    y     = NULL
  )
```

```{r stage1-parcel-table}
stage1 |>
  mutate(
    parcel_status = if_else(no_parcel_found, "No parcel found", "Parcel found"),
    flagged       = stage1_prob_correct < STAGE1_THRESHOLD | no_parcel_found
  ) |>
  group_by(parcel_status) |>
  summarise(
    n            = n(),
    n_flagged    = sum(flagged),
    pct_flagged  = mean(flagged),
    median_score = median(stage1_prob_correct),
    mean_score   = mean(stage1_prob_correct),
    .groups = "drop"
  ) |>
  gt() |>
  tab_header(title = "Stage 1 Results by Parcel Found Status") |>
  fmt_percent(columns = pct_flagged, decimals = 1) |>
  fmt_number(columns = c(median_score, mean_score), decimals = 3) |>
  fmt_integer(columns = c(n, n_flagged)) |>
  cols_label(
    parcel_status = "Parcel Status",
    n             = "N Plants",
    n_flagged     = "N Flagged",
    pct_flagged   = "% Flagged",
    median_score  = "Median Score",
    mean_score    = "Mean Score"
  ) |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  opt_table_font(font = google_font("IBM Plex Sans"))
```

## Trigger Reason Breakdown

The `trigger_reason` field records which Stage 1 condition sent a plant to Stage 2.

```{r trigger-reason}
#| fig-cap: "Breakdown of `trigger_reason` for all Stage 1-flagged plants."

stage1 |>
  filter(!is.na(trigger_reason)) |>
  count(trigger_reason, sort = TRUE) |>
  mutate(trigger_reason = fct_reorder(trigger_reason, n)) |>
  ggplot(aes(x = n, y = trigger_reason)) +
  geom_col(fill = epa_blue, width = 0.65) +
  geom_text(aes(label = comma(n)), hjust = -0.15, size = 3.2, colour = epa_grey) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Stage 1 Trigger Reasons",
    x     = "Number of plants",
    y     = NULL
  )
```

---

# Tier Breakdown

Plants are assigned to one of six tiers based on Stage 1 confidence and Stage 2
candidate quality. `tier_label` is derived from the `tier` column using the
mapping defined in the setup chunk.

```{r tier-bar}
#| fig-cap: "Count of CWNS plants assigned to each correction tier."

summary_df |>
  count(tier_label) |>
  mutate(tier_label = fct_reorder(tier_label, n)) |>
  ggplot(aes(x = n, y = tier_label, fill = tier_label)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = comma(n)), hjust = -0.15, size = 3.2, colour = epa_grey) +
  scale_fill_manual(values = tier_colors) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Plant Tier Assignment",
    x     = "Number of plants",
    y     = NULL
  )
```

```{r tier-table}
summary_df |>
  count(tier_label) |>
  mutate(
    pct     = n / sum(n),
    cum_n   = cumsum(n),
    cum_pct = cumsum(pct)
  ) |>
  gt() |>
  tab_header(title = "Tier Assignment Summary") |>
  fmt_percent(columns = c(pct, cum_pct), decimals = 1) |>
  fmt_integer(columns = c(n, cum_n)) |>
  cols_label(
    tier_label = "Tier",
    n          = "N",
    pct        = "%",
    cum_n      = "Cumulative N",
    cum_pct    = "Cumulative %"
  ) |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  data_color(
    columns = pct,
    palette = c("white", epa_blue),
    domain  = c(0, 1)
  ) |>
  opt_table_font(font = google_font("IBM Plex Sans"))
```

---

# Stage 2 — Candidate Parcel Scoring

## Score Distribution

Stage 2 scores each candidate parcel with `stage2_prob_correct`. The threshold
of **0.057** is intentionally low — it reflects the 1:50 hard-negative sampling
ratio used during training.

```{r stage2-hist}
#| fig-cap: "Distribution of `stage2_prob_correct` across all 318,011 candidate
#|   parcels. The low threshold is expected given the 1:50 hard-negative training regime."

stage2 |>
  ggplot(aes(x = stage2_prob_correct)) +
  geom_histogram(bins = 100, fill = epa_blue, colour = "white", linewidth = 0.1) +
  geom_vline(xintercept = STAGE2_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.9) +
  annotate("text", x = STAGE2_THRESHOLD + 0.003, y = Inf,
           label = glue("Threshold\n{STAGE2_THRESHOLD}"),
           vjust = 1.3, hjust = 0, size = 3.2, colour = epa_red) +
  scale_x_continuous(labels = percent_format(1)) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Stage 2 Candidate Parcel Score Distribution",
    subtitle = glue("{comma(n_cands)} candidates across {comma(n_stage2)} flagged plants"),
    x        = "stage2_prob_correct",
    y        = "Number of candidate parcels",
    caption  = "Low threshold expected due to 1:50 hard-negative sampling during training."
  )
```

## Candidates per Plant

```{r cands-per-plant}
#| fig-cap: "Distribution of candidate parcel count per flagged plant. Candidates
#|   are drawn from k=9 H3 rings (~5 km radius from reported location)."

cands_per_plant <- stage2 |>
  count(CWNS_ID, name = "n_candidates")

p1 <- cands_per_plant |>
  ggplot(aes(x = n_candidates)) +
  geom_histogram(binwidth = 1, fill = epa_blue, colour = "white", linewidth = 0.1) +
  coord_cartesian(xlim = c(0,50))+
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  labs(title = "Candidates per Plant", x = "Number of candidate parcels", y = "Number of plants")

p2 <- cands_per_plant |>
  ggplot(aes(y = n_candidates)) +
  geom_boxplot(fill = epa_blue, alpha = 0.6, width = 0.4,
               outlier.colour = epa_red, outlier.alpha = 0.3) +
  scale_y_continuous(labels = comma) +
  labs(y = "Candidates per plant", x = NULL) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p1 + p2 + plot_layout(widths = c(3, 1))
```

## Best Score per Plant (`top_stage2_prob`)

The summary CSV carries the top-ranked candidate per plant as `top_stage2_prob`.
This is the key decision variable for determining whether a plant receives a
correction.

```{r top-score}
#| fig-cap: "Distribution of `top_stage2_prob` for flagged plants. Green = accepted
#|   candidate (above threshold); grey = no strong candidate found."

best_per_plant <- summary_df |>
  filter(!is.na(top_stage2_prob)) |>
  mutate(has_candidate = top_stage2_prob >= STAGE2_THRESHOLD)

n_with_cand <- sum(best_per_plant$has_candidate)

best_per_plant |>
  ggplot(aes(x = top_stage2_prob, fill = has_candidate)) +
  geom_histogram(bins = 80, colour = "white", linewidth = 0.1) +
  geom_vline(xintercept = STAGE2_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.9) +
  scale_fill_manual(
    values = c("TRUE" = epa_green, "FALSE" = epa_grey),
    labels = c("TRUE" = "Candidate accepted", "FALSE" = "No strong candidate")
  ) +
  scale_x_continuous(labels = percent_format(1)) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Best Stage 2 Score per Flagged Plant",
    subtitle = glue(
      "{comma(n_with_cand)} of {comma(n_stage2)} flagged plants ",
      "({percent(n_with_cand / n_stage2, 0.1)}) have an accepted candidate"
    ),
    x    = "top_stage2_prob",
    y    = "Number of plants",
    fill = NULL
  )
```

## Proximity of Accepted Candidates

Stage 2 records `within_1km` and `within_5km` flags for each candidate, carried
through to the summary for the top candidate.

```{r proximity-breakdown}
#| fig-cap: "Proximity breakdown of accepted Stage 2 candidates using `within_1km`
#|   and `within_5km` flags."

summary_df |>
  filter(!is.na(top_stage2_prob), top_stage2_prob >= STAGE2_THRESHOLD) |>
  mutate(proximity = case_when(
    as.logical(within_1km)                          ~ "Within 1 km",
    as.logical(within_5km) & !as.logical(within_1km) ~ "1 to 5 km",
    !as.logical(within_5km)                         ~ "Beyond 5 km"
  )) |>
  count(proximity) |>
  mutate(
    proximity = factor(proximity, levels = c("Within 1 km", "1 to 5 km", "Beyond 5 km")),
    pct       = n / sum(n)
  ) |>
  ggplot(aes(x = n, y = proximity, fill = proximity)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = glue("{comma(n)} ({percent(pct, 0.1)})")),
            hjust = -0.1, size = 3.2, colour = epa_grey) +
  scale_fill_manual(values = c(epa_green, epa_gold, epa_red)) +
  scale_x_continuous(labels = comma, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Proximity of Accepted Candidates to Reported Location",
    x     = "Number of plants",
    y     = NULL
  )
```

---

# Correction Distance Analysis

`top_distance_m` records the distance in metres from the reported plant location
to the top-ranked Stage 2 candidate parcel.

## Overall Distance Distribution

```{r distance-dist}
#| fig-cap: "Distribution of `top_distance_m` for plants with an accepted correction.
#|   Log scale used due to the long right tail."

corrected <- summary_df |>
  filter(!is.na(top_distance_m), top_stage2_prob >= STAGE2_THRESHOLD)

med_dist  <- median(corrected$top_distance_m)
mean_dist <- mean(corrected$top_distance_m)

corrected |>
  ggplot(aes(x = top_distance_m)) +
  geom_histogram(bins = 80, fill = epa_blue, colour = "white", linewidth = 0.1) +
  geom_vline(xintercept = med_dist,  colour = epa_green, linetype = "dashed", linewidth = 0.9) +
  geom_vline(xintercept = mean_dist, colour = epa_gold,  linetype = "dashed", linewidth = 0.9) +
  annotate("text", x = med_dist,  y = Inf,
           label = glue("Median\n{comma(med_dist, 1)} m"),
           vjust = 1.3, hjust = -0.1, size = 3, colour = epa_green) +
  annotate("text", x = mean_dist, y = Inf,
           label = glue("Mean\n{comma(mean_dist, 1)} m"),
           vjust = 1.3, hjust = -0.1, size = 3, colour = epa_gold) +
  scale_x_log10(labels = label_comma()) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Correction Distance Distribution",
    subtitle = glue("{comma(nrow(corrected))} plants with accepted correction"),
    x        = "top_distance_m (log scale)",
    y        = "Number of plants"
  )
```

## Distance by Tier

```{r distance-by-tier}
#| fig-cap: "Correction distance by tier. The T1 tier boundary at 5,000 m
#|   is visible as the split between 'nearby' and 'distant' tiers."

summary_df |>
  filter(!is.na(top_distance_m), top_stage2_prob >= STAGE2_THRESHOLD) |>
  ggplot(aes(x = top_distance_m, y = tier_label, fill = tier_label)) +
  geom_density_ridges(alpha = 0.75, quantile_lines = TRUE,
                      quantiles = 0.5, colour = "white", scale = 0.9) +
  scale_fill_manual(values = tier_colors, guide = "none") +
  scale_x_log10(labels = label_comma()) +
  labs(
    title = "Correction Distance by Tier",
    x     = "top_distance_m (log scale)",
    y     = NULL
  )
```

## Cumulative Correction Distance

```{r ecdf-distance}
#| fig-cap: "ECDF of correction distances. Dotted lines mark common thresholds."

distance_cuts <- c(100, 250, 500, 1000, 5000)

corrected |>
  ggplot(aes(x = top_distance_m)) +
  stat_ecdf(colour = epa_blue, linewidth = 1.1) +
  geom_vline(xintercept = distance_cuts, colour = epa_grey,
             linetype = "dotted", linewidth = 0.5) +
  annotate("text",
           x     = distance_cuts, y = 0.02,
           label = paste0(distance_cuts / 1000, " km"),
           size  = 2.8, colour = epa_grey, hjust = -0.1) +
  scale_x_log10(labels = label_comma()) +
  scale_y_continuous(labels = percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(
    title = "Cumulative Distribution of Correction Distances",
    x     = "top_distance_m (log scale)",
    y     = "Cumulative proportion of corrected plants"
  )
```

```{r distance-summary-table}
corrected |>
  summarise(
    N              = n(),
    `Min (m)`      = min(top_distance_m),
    `P10 (m)`      = quantile(top_distance_m, 0.10),
    `Median (m)`   = median(top_distance_m),
    `Mean (m)`     = mean(top_distance_m),
    `P90 (m)`      = quantile(top_distance_m, 0.90),
    `Max (m)`      = max(top_distance_m),
    `% < 500 m`    = mean(top_distance_m < 500),
    `% 500 m-5 km` = mean(top_distance_m >= 500 & top_distance_m < 5000),
    `% >= 5 km`    = mean(top_distance_m >= 5000)
  ) |>
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") |>
  mutate(
    formatted = case_when(
      str_detect(Statistic, "^%") ~ percent(Value, 0.1),
      Statistic == "N"            ~ comma(Value, 1),
      TRUE                        ~ comma(Value, 1)
    )
  ) |>
  select(Statistic, Value = formatted) |>
  gt() |>
  tab_header(title = "Correction Distance Summary Statistics") |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  opt_row_striping() |>
  opt_table_font(font = google_font("IBM Plex Sans"))
```

---

# High-Confidence Corrections — Spot-Check

Top 25 nearby corrections (Tier 1 nearby or Tier 2 corrected) ranked by
`top_stage2_prob`. The `top_ll_uuid` column identifies the accepted parcel in
the Regrid dataset.

```{r spot-check}
summary_df |>
  filter(tier %in% c("Tier1_nearby_correction", "Tier2_no_parcel_corrected")) |>
  arrange(desc(top_stage2_prob)) |>
  slice_head(n = 25) |>
  select(
    CWNS_ID,
    FACILITY_NAME,
    STATE_CODE,
    tier_label,
    stage1_prob_correct,
    top_stage2_prob,
    top_distance_m,
    top_ll_uuid
  ) |>
  gt() |>
  tab_header(
    title    = "Top 25 High-Confidence Corrections",
    subtitle = "Ranked by top_stage2_prob — nearby corrections only"
  ) |>
  fmt_number(columns = c(stage1_prob_correct, top_stage2_prob), decimals = 3) |>
  fmt_number(columns = top_distance_m, decimals = 0) |>
  cols_label(
    CWNS_ID             = "CWNS ID",
    FACILITY_NAME       = "Facility",
    STATE_CODE          = "State",
    tier_label          = "Tier",
    stage1_prob_correct = "Stage 1",
    top_stage2_prob     = "Stage 2",
    top_distance_m      = "Distance (m)",
    top_ll_uuid         = "Parcel UUID"
  ) |>
  data_color(
    columns = top_stage2_prob,
    palette = c("white", epa_green),
    domain  = c(0, 1)
  ) |>
  data_color(
    columns = stage1_prob_correct,
    palette = c(epa_red, "white"),
    domain  = c(0, 1)
  ) |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  opt_table_font(font = google_font("IBM Plex Sans")) |>
  tab_options(table.font.size = "small")
```

---

# Distant Corrections — Manual Review Candidates

Tier 1 distant corrections (beyond 5 km per `within_5km = FALSE`) require manual
QA before applying. Ranked by `top_stage2_prob` to prioritise the most confident
predictions.

```{r distant-table}
summary_df |>
  filter(tier == "Tier1_distant_correction") |>
  arrange(desc(top_stage2_prob)) |>
  select(
    CWNS_ID,
    FACILITY_NAME,
    STATE_CODE,
    stage1_prob_correct,
    top_stage2_prob,
    top_distance_m,
    top_ll_uuid
  ) |>
  slice_head(n = 30) |>
  gt() |>
  tab_header(
    title    = "Distant Corrections (beyond 5 km)",
    subtitle = "Top 30 by top_stage2_prob — prioritise these for manual GIS review"
  ) |>
  fmt_number(columns = c(stage1_prob_correct, top_stage2_prob), decimals = 3) |>
  fmt_number(columns = top_distance_m, decimals = 0) |>
  cols_label(
    CWNS_ID             = "CWNS ID",
    FACILITY_NAME       = "Facility",
    STATE_CODE          = "State",
    stage1_prob_correct = "Stage 1",
    top_stage2_prob     = "Stage 2",
    top_distance_m      = "Distance (m)",
    top_ll_uuid         = "Parcel UUID"
  ) |>
  data_color(
    columns = top_distance_m,
    palette = c("white", epa_red),
    domain  = NULL
  ) |>
  tab_style(
    style = cell_text(weight = "bold", color = epa_blue),
    locations = cells_column_labels()
  ) |>
  opt_table_font(font = google_font("IBM Plex Sans")) |>
  tab_options(table.font.size = "small")
```

---

# Geographic Distribution

## State-Level Summary

```{r state-summary}
#| fig-height: 7
#| fig-cap: "Proportion of plants flagged by Stage 1 by state (`STATE_CODE`). High
#|   flag rates may reflect parcel coverage gaps or systematic geocoding errors in CWNS."

state_summary <- summary_df |>
  group_by(STATE_CODE) |>
  summarise(
    n_plants      = n(),
    n_flagged     = sum(tier != "Tier3_likely_correct"),
    pct_flagged   = n_flagged / n_plants,
    n_corrected   = sum(tier %in% tier_raw_corrected),
    pct_corrected = n_corrected / n_plants,
    .groups = "drop"
  ) |>
  arrange(desc(pct_flagged))

state_summary |>
  slice_head(n = 30) |>
  mutate(STATE_CODE = fct_reorder(STATE_CODE, pct_flagged)) |>
  ggplot(aes(x = pct_flagged, y = STATE_CODE)) +
  geom_col(fill = epa_blue, width = 0.7) +
  geom_col(aes(x = pct_corrected), fill = epa_green, width = 0.7, alpha = 0.85) +
  scale_x_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.05))) +
  labs(
    title    = "Top 30 States by Flag Rate",
    subtitle = "Blue = flagged · Green overlay = successfully corrected",
    x        = "Proportion of plants",
    y        = "State"
  )
```

## Tier Composition by State

```{r tier-by-state}
#| fig-height: 8
#| fig-cap: "Tier composition (proportion) for the 20 most-flagged states."

top20_states <- state_summary |>
  slice_max(n_flagged, n = 20) |>
  pull(STATE_CODE)

summary_df |>
  filter(STATE_CODE %in% top20_states) |>
  count(STATE_CODE, tier_label) |>
  group_by(STATE_CODE) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  mutate(STATE_CODE = factor(STATE_CODE, levels = rev(top20_states))) |>
  ggplot(aes(x = prop, y = STATE_CODE, fill = tier_label)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = tier_colors) +
  scale_x_continuous(labels = percent_format()) +
  labs(
    title = "Tier Composition — Top 20 States by Flagged Plants",
    x     = "Proportion of plants",
    y     = NULL,
    fill  = NULL
  ) +
  theme(legend.position = "right")
```

## Rural vs. Urban Flag Rates

```{r rural-urban}
#| fig-cap: "Stage 1 score distributions by `is_rural`. Rural plants may have
#|   sparser parcel data, affecting both flag rates and correction availability."

summary_df |>
  mutate(setting = if_else(is_rural, "Rural", "Urban/suburban")) |>
  ggplot(aes(x = stage1_prob_correct, y = setting, fill = setting)) +
  geom_density_ridges(alpha = 0.75, quantile_lines = TRUE,
                      quantiles = 0.5, colour = "white") +
  geom_vline(xintercept = STAGE1_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.8) +
  scale_fill_manual(values = c("Rural" = epa_gold, "Urban/suburban" = epa_blue),
                    guide = "none") +
  scale_x_continuous(labels = percent_format(1), limits = c(0, 1)) +
  labs(
    title = "Stage 1 Scores by Rural/Urban Setting",
    x     = "stage1_prob_correct",
    y     = NULL
  )
```

## Population Served vs. Confidence Score

```{r pop-vs-score}
#| fig-cap: "Relationship between `pop_served` and `stage1_prob_correct`.
#|   Larger plants may have more reliable reported coordinates."

summary_df |>
  filter(!is.na(pop_served), pop_served > 0) |>
  ggplot(aes(x = pop_served, y = stage1_prob_correct,
             colour = tier == "Tier3_likely_correct")) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_hline(yintercept = STAGE1_THRESHOLD, colour = epa_red,
             linetype = "dashed", linewidth = 0.7) +
  scale_x_log10(labels = label_comma()) +
  scale_colour_manual(
    values = c("TRUE" = epa_green, "FALSE" = epa_red),
    labels = c("TRUE" = "Likely correct", "FALSE" = "Flagged"),
    name   = NULL
  ) +
  labs(
    title = "Population Served vs. Stage 1 Confidence",
    x     = "Population served (log scale)",
    y     = "stage1_prob_correct"
  )
```

---

# Geopackage Layer Review

The `inference_results.gpkg` file contains three spatial layers for visual
inspection in QGIS or R. The `eval` condition below skips this section if the
file is not present.

```{r gpkg-layers}
#| eval: !expr file.exists(path_gpkg)

layers <- st_layers(path_gpkg)$name
cat("Layers found in geopackage:\n")
print(layers)
```

```{r gpkg-preview}
#| eval: !expr file.exists(path_gpkg)
#| fig-cap: "Sample of up to 500 corrected plants: reported (red), accepted
#|   candidate (green), and correction vectors (grey lines)."

# Adjust layer indices below if the layer order differs
reported   <- st_read(path_gpkg, layer = layers[1], quiet = TRUE)
candidates <- st_read(path_gpkg, layer = layers[2], quiet = TRUE)
lines      <- st_read(path_gpkg, layer = layers[3], quiet = TRUE)

corrected_ids <- summary_df |>
  filter(tier %in% tier_raw_corrected) |>
  pull(CWNS_ID)

set.seed(42)
sample_ids <- sample(corrected_ids, min(500, length(corrected_ids)))

# Assumes CWNS_ID is the join key in the geopackage layers; adjust if needed
rep_samp  <- reported   |> filter(CWNS_ID %in% sample_ids)
cand_samp <- candidates |> filter(CWNS_ID %in% sample_ids)
line_samp <- lines      |> filter(CWNS_ID %in% sample_ids)

ggplot() +
  geom_sf(data = line_samp,  colour = epa_grey,  linewidth = 0.3, alpha = 0.5) +
  geom_sf(data = cand_samp,  colour = epa_green, size = 1.2, shape = 16, alpha = 0.7) +
  geom_sf(data = rep_samp,   colour = epa_red,   size = 1.2, shape = 16, alpha = 0.7) +
  labs(
    title    = "Reported (red) to Corrected (green) Location Sample",
    subtitle = glue("Random sample of {length(sample_ids)} corrected plants"),
    caption  = "Lines connect reported to accepted candidate location."
  ) +
  theme(axis.text = element_text(size = 7))
```

---

# Summary & Recommendations

```{r tier-counts}
#| echo: false
n_t1_near <- sum(summary_df$tier == "Tier1_nearby_correction")
n_t1_dist <- sum(summary_df$tier == "Tier1_distant_correction")
n_t1_none <- sum(summary_df$tier == "Tier1_no_candidate")
n_t2_corr <- sum(summary_df$tier == "Tier2_no_parcel_corrected")
n_t2_none <- sum(summary_df$tier == "Tier2_no_parcel_uncorrected")
n_t3      <- sum(summary_df$tier == "Tier3_likely_correct")
```

## Actionable Tier Summary

| Priority | Tier | N | Recommended Action |
|----------|------|---|--------------------|
| Done | T3 — Likely correct | `r comma(n_t3)` | Accept reported location; no action needed |
| Auto-apply | T1 — Nearby correction | `r comma(n_t1_near)` | Apply `top_ll_uuid` candidate; spot-check the high-distance tail |
| Auto-apply | T2 — No parcel, corrected | `r comma(n_t2_corr)` | Apply `top_ll_uuid` candidate; confirm parcel is plausible |
| Manual QA | T1 — Distant correction | `r comma(n_t1_dist)` | Sort by `top_stage2_prob` desc; verify each in QGIS before applying |
| Manual QA | T1 — No candidate | `r comma(n_t1_none)` | Cross-reference CWNS address fields; consider geocoding fallback |
| Flag only | T2 — No parcel, uncorrected | `r comma(n_t2_none)` | Flag as unresolved; likely reflects Regrid coverage gaps |

## Next Steps

1. **Apply automatic corrections** for T1-nearby and T2-corrected tiers. Export a
   correction table keyed by `CWNS_ID` with columns `top_ll_uuid`,
   `top_distance_m`, and `tier`.

2. **Manual QA queue** — export T1-distant plants sorted descending by
   `top_stage2_prob`. Step through correction lines in `inference_results.gpkg`
   using QGIS to visually assess each candidate before accepting.

3. **T1-no-candidate investigation** — cross-reference `CWNS_ID` against CWNS
   address fields. A Stage 3 address-geocoding pass (Census Bureau geocoder or
   TIGER/Line) could recover some of these cases.

4. **T2-uncorrected coverage audit** — check whether `Tier2_no_parcel_uncorrected`
   plants cluster by `STATE_CODE`. Regrid parcel coverage is uneven across states
   and may explain the pattern.

5. **Model retraining** — verified corrections from this run can be added to the
   Stage 2 training set (currently 203 positives). Target 400-500 verified
   positives before the next iteration to meaningfully improve recall on distant
   corrections.

---

# Session Info

```{r session-info}
sessioninfo::session_info()
```

```{r rename-output}
#| echo: false
#| message: false
# ── Rename output file to reflect the model version ───────────────────────────
# Quarto does not support !expr in output-file, so we rename post-render.
# This chunk runs last and renames the default .html output to
#   cwns_location_correction_report_Version_N.html
# in the same directory as the .qmd file.

.qmd_dir     <- here::here()
.default_out <- file.path(.qmd_dir, "cwns_location_correction_report.html")
.target_out  <- file.path(.qmd_dir,
                           paste0("cwns_location_correction_report_",
                                  version_label, ".html"))

if (file.exists(.default_out) && .default_out != .target_out) {
  file.rename(.default_out, .target_out)
  message("Output renamed to: ", basename(.target_out))
} else if (file.exists(.target_out)) {
  message("Output already at: ", basename(.target_out))
} else {
  message("Note: default output not found for renaming (",
          basename(.default_out), ") — file may have been named differently.")
}
```
